PineAPPL:\\ EW and mixed QCD--EW corrections for PDF fits

### 2020-08-06 paper/sec-results.tex
Updated ERN funding
added some comments in results. Removed commented tex in doublecounting

\paragraph{Top-quark pair production.}
We select the distributions, single-differential in either the transverse
momentum of the top quark, $p_\mathrm{T}^\mathrm{t}$, or the invariant mass of the top-quark
pair, $m_{\mathrm{t}\bar{\mathrm{t}}}$, measured by the ATLAS and CMS experiments at a centre-of-mass
energy of \SI{8}{\tera\electronvolt}~\cite{Aad:2015mbv,Khachatryan:2015oqa}. These measurements have
been extensively studied in the context of PDF fits in
refs.~\cite{Czakon:2016olj,Bailey:2019yze,Amoroso:2020lgh,Kadir:2020yml}, and
included by default in the CT18~\cite{Hou:2019efy} analysis.
Because EW corrections are significantly smaller for distributions differential
in the rapidity of either the top quark or the top-quark
pair~\cite{Czakon:2017wor}, these distributions were preferred for inclusion
in the NNPDF3.1 analysis~\cite{Ball:2017nwa}. The process receives
pure QCD contributions at LO, $\mathcal{O}(\alphas^2)$, and
at NLO, $\mathcal{O}(\alphas^3)$. They make up our NLO QCD
computation. The NLO QCD+EW computation includes the $\mathcal{O}(\alphas^2)$
and $\mathcal{O}(\alphas^3)$ QCD contributions, the LO contribution
$\mathcal{O}(\alphas\alpha)$ and the NLO contribution
$\mathcal{O}(\alphas^2\alpha)$.
We do not consider the LO contribution $\mathcal{O}(\alpha^2)$ nor the
NLO contribution $\mathcal{O}(\alphas\alpha^2)$.
EW corrections for this process
were computed in refs.~\cite{Bernreuther:2010ny,Hollik:2011ps,Kuhn:2011ri,Bernreuther:2012sx,Pagani:2016caq,Czakon:2017wor,Czakon:2017lgo,Czakon:2017mmr,Czakon:2019bcq,Czakon:2019txp}. The process receives contributions from
6 (36) parton luminosities at NLO QCD (NLO QCD+EW),
see appendix~\ref{app:lumis} for details.

@@

The values of the relevant physical parameters are chosen as
\begin{equation}
\begin{aligned}
M_\mathrm{W} &= \SI{80}{\giga\electronvolt} \text{,} \quad &
M_\mathrm{Z} &= \SI{91.176}{\giga\electronvolt} \text{,} \quad &
m_\mathrm{t} &= \SI{172.5}{\giga\electronvolt} \text{,} \quad \\
\Gamma_\mathrm{W} &= \SI{2.09}{\giga\electronvolt} \text{,} &
\Gamma_\mathrm{Z} &= \SI{2.50}{\giga\electronvolt} \text{,}
\end{aligned}
\label{eq:parameters}
\end{equation}
where $M_\mathrm{W}$, $M_\mathrm{Z}$ and $m_\mathrm{t}$ are the values of the W-boson, Z-boson, and
top quark masses, respectively, and $\Gamma_\mathrm{W}$ and $\Gamma_\mathrm{Z}$ are the width of
the W boson and of the Z boson, respectively. The value of the strong
coupling is chosen consistently with the PDF set, $\alphas(M_\mathrm{Z})=0.118$.
When relevant, final-state photons and massless charged fermions (leptons and light quarks) are recombined together 
if they satisfy the condition $\Delta R_{f \gamma}<0.1$, where $\Delta R_{f \gamma}$
is the fermion-photon cone radius. In this case the
sum of their momenta is assigned to the charged fermion, and the photon is removed
from the event. Kinematics observables and cuts are defined starting from recombined momenta. If we were interested also in jet-related observables,
photons surviving the recombination wouldl have to be clustered together with coloured
partons.\footnote{For issues related to the definition of jets in presence of EW corrections,
    see refs.~\cite{Frederix:2016ost,Denner:2019zfp}}

@@

%-------------------------------------------------------------------------------
\begin{figure}[!t]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_ATLASZHIGHMASS49FB}%
    \includegraphics[width=0.5\textwidth]{figures/pineappl_ATLASZHIGHMASS49FB_lowstat}
    \caption{Validation and test of the \textsc{PineAPPL} grid for the ATLAS
      high-mass Drell--Yan lepton pair measurement in the high-mass region at
      a centre-of-mass energy of \SI{7}{\tera\electronvolt}~\cite{Aad:2013iua}. The first panel
      displays the relative difference (in per mille) between the {\sc PineAPPL}
      and the MC results for the central, upper and lower scale choices,
      for theories accurate to both NLO QCD and NLO QCD+EW. The second, third
      and fourth panels present the theoretical predictions, accurate to either
      NLO QCD or NLO QCD+EW, always normalised to the former; on top of the
      theoretical predictions, the PDF uncertainty, the scale uncertainty and
      the Monte Carlo uncertainty are displayed in turn. The relative
      uncertainty of the experimental data is also shown for comparison.}
    \label{fig:atlaszhighmass49fb}
\end{figure}
%-------------------------------------------------------------------------------

@@

The process receives its leading contribution from the $\mathrm{gg}$ channel,
which varies between 81\% and 61\% (76\% and 83\%) of the $p_\mathrm{T}^\mathrm{t}$
($m_{\mathrm{t}\bar{\mathrm{t}}}$) differential cross section as the value of the
transverse momentum of the top quark (the invariant mass of the top-quark pair)
increases in the bin range; $\mathrm{\gamma}\mathrm{g}$ scattering
correspondingly accounts for about 0.5\% -- 1\% (0.5\% -- 0.7\%) of the cross
section; the contribution from other PI parton luminosities is comparatively
negligible. Overall, the EW corrections suppress the $p_\mathrm{T}^\mathrm{t}$
($m_{\mathrm{t}\bar{\mathrm{t}}}$) distribution by about 0.2\% -- 3.5\%
(0.5\% -- 0.2\%) for increasing values of $p_\mathrm{T}^\mathrm{t}$
($m_{\mathrm{t}\bar{\mathrm{t}}}$), except in the first bin of the
$p_\mathrm{T}^\mathrm{t}$ distribution, where they enhance the cross section by
about 1\%. The size of these shifts, however, remains always significantly
smaller than the data uncertainty.\footnote{In
  figure~\ref{fig:atlastop} the data uncertainty corresponds to the ATLAS
  measurement~\cite{Aad:2015auj}. Similar considerations apply also for the CMS
  measurement~\cite{Khachatryan:2015oaa}.}
As a consequence, we anticipate that the more accurate NLO QCD+EW theory is
likely to be easily accommodated by the large data uncertainty, should the data
be fitted with the inclusion of EW corrections; PDFs will possibly not become
more precise.

>>>>

\paragraph{Top-quark pair production.}
We select the distributions, single-differential in either the transverse
momentum of the top quark, $p_\mathrm{T}^\mathrm{t}$, or the invariant mass of the top-quark
pair, $m_{\mathrm{t}\bar{\mathrm{t}}}$, measured by the ATLAS and CMS experiments at a centre-of-mass
energy of \SI{8}{\tera\electronvolt}~\cite{Aad:2015mbv,Khachatryan:2015oqa}. These measurements have
been extensively studied in the context of PDF fits in
refs.~\cite{Czakon:2016olj,Bailey:2019yze,Amoroso:2020lgh,Kadir:2020yml} (see
also refs.~\cite{Pagani:2016caq,Czakon:2017wor} for studies related to the photon density) and
included by default in the CT18~\cite{Hou:2019efy} analysis.
Because EW corrections are significantly smaller for distributions differential
in the rapidity of either the top quark or the top-quark
pair~\cite{Czakon:2017wor}, these distributions were preferred for inclusion
in the NNPDF3.1 analysis~\cite{Ball:2017nwa}. The process receives
pure QCD contributions at LO, $\mathcal{O}(\alphas^2)$, and
at NLO, $\mathcal{O}(\alphas^3)$. They make up our NLO QCD
computation. The NLO QCD+EW computation includes the $\mathcal{O}(\alphas^2)$
and $\mathcal{O}(\alphas^3)$ QCD contributions, the LO contribution
$\mathcal{O}(\alphas\alpha)$ and the NLO contribution
$\mathcal{O}(\alphas^2\alpha)$.
We do not consider the LO contribution $\mathcal{O}(\alpha^2)$ nor the
NLO contributions $\mathcal{O}(\alphas\alpha^2)$ and $\mathcal{O}(\alpha^3)$, which have been shown to be negligible~\cite{Czakon:2017wor,Frederix:2018nkq}.
EW corrections for this process
were computed in refs.~\cite{Bernreuther:2010ny,Hollik:2011ps,Kuhn:2011ri,Bernreuther:2012sx,Pagani:2016caq,Czakon:2017wor,Czakon:2017lgo,Czakon:2017mmr,Czakon:2019bcq,Czakon:2019txp}. The process receives contributions from
6 (36) parton luminosities at NLO QCD (NLO QCD+EW),
see appendix~\ref{app:lumis} for details.

@@

The values of the relevant physical parameters are chosen as
\begin{equation}
\begin{aligned}
M_\mathrm{W} &= \SI{80}{\giga\electronvolt} \text{,} \quad &
M_\mathrm{Z} &= \SI{91.176}{\giga\electronvolt} \text{,} \quad &
m_\mathrm{t} &= \SI{172.5}{\giga\electronvolt} \text{,} \quad \\
\Gamma_\mathrm{W} &= \SI{2.09}{\giga\electronvolt} \text{,} &
\Gamma_\mathrm{Z} &= \SI{2.50}{\giga\electronvolt} \text{,}
\end{aligned}
\label{eq:parameters}
\end{equation}
where $M_\mathrm{W}$, $M_\mathrm{Z}$ and $m_\mathrm{t}$ are the values of the W-boson, Z-boson, and
top quark masses, respectively, and $\Gamma_\mathrm{W}$ and $\Gamma_\mathrm{Z}$ are the width of
the W boson and of the Z boson, respectively. The value of the strong
coupling is chosen consistently with the PDF set, $\alphas(M_\mathrm{Z})=0.118$.
When relevant, final-state photons and massless charged fermions (leptons and light quarks) are recombined together 
if they satisfy the condition $\Delta R_{f \gamma}<0.1$, where $\Delta R_{f \gamma}$
is the fermion-photon cone radius. In this case the
sum of their momenta is assigned to the charged fermion, and the photon is removed
from the event. Kinematics observables and cuts are defined starting from recombined momenta. If we were interested also in jet-related observables,
photons surviving the recombination wouldl have to be clustered together with coloured
partons.\footnote{For issues related to the definition of jets in presence of EW corrections,
see refs.~\cite{Frederix:2016ost,Denner:2019zfp}.} Finally, although contributions corresponding to the radiation
of a heavy boson are formally of the same perturbative order as the EW corrections, they are not included in our
computations. In fact, while nothing prevents to include these contributions \emph{a posteriori}, as they are finite,
their impact is either smaller than the one of ``standard'' EW corrections, or anyway negligible with respect to
the total cross section (see refs.~\cite{Frixione:2014qaa,Frixione:2015zaa,Pagani:2016caq} for some process-specific cases {\bf MZ OTHERS???}).

@@

%-------------------------------------------------------------------------------
\begin{figure}[!t]
    \centering
    \includegraphics[width=0.5\textwidth]{figures/pineappl_ATLASZHIGHMASS49FB}%
    \includegraphics[width=0.5\textwidth]{figures/pineappl_ATLASZHIGHMASS49FB_lowstat}\\
    \caption{Validation and test of the \textsc{PineAPPL} grid for the ATLAS
      high-mass Drell--Yan lepton pair measurement in the high-mass region at
      a centre-of-mass energy of \SI{7}{\tera\electronvolt}~\cite{Aad:2013iua}. The first panel
      displays the relative difference (in per mille) between the {\sc PineAPPL}
      and the MC results for the central, upper and lower scale choices,
      for theories accurate to both NLO QCD and NLO QCD+EW. The second, third
      and fourth panels present the theoretical predictions, accurate to either
      NLO QCD or NLO QCD+EW, always normalised to the former; on top of the
      theoretical predictions, the PDF uncertainty, the scale uncertainty and
      the Monte Carlo uncertainty are displayed in turn. The relative
      uncertainty of the experimental data is also shown for comparison.}
    \label{fig:atlaszhighmass49fb}
\end{figure}
%-------------------------------------------------------------------------------

@@

The process receives its leading contribution from the $\mathrm{gg}$ channel,
which varies between 81\% and 61\% (76\% and 83\%) of the $p_\mathrm{T}^\mathrm{t}$
($m_{\mathrm{t}\bar{\mathrm{t}}}$) differential cross section as the value of the
transverse momentum of the top quark (the invariant mass of the top-quark pair)
increases in the bin range; $\mathrm{\gamma}\mathrm{g}$ scattering
correspondingly accounts for about 0.5\% -- 1\% (0.5\% -- 0.7\%) of the cross
section, and is almost entirely {\bf MZ CHECK!!!} due to the LO contribution at $\mathcal{O}(\alphas\alpha)$; the contribution from other PI parton luminosities is comparatively
negligible. Overall, the EW corrections suppress the $p_\mathrm{T}^\mathrm{t}$
($m_{\mathrm{t}\bar{\mathrm{t}}}$) distribution by about 0.2\% -- 3.5\%
(0.5\% -- 0.2\%) for increasing values of $p_\mathrm{T}^\mathrm{t}$
($m_{\mathrm{t}\bar{\mathrm{t}}}$), except in the first bin of the
$p_\mathrm{T}^\mathrm{t}$ distribution, where they enhance the cross section by
about 1\%. The size of these shifts, however, remains always significantly
smaller than the data uncertainty.\footnote{In
  figure~\ref{fig:atlastop} the data uncertainty corresponds to the ATLAS
  measurement~\cite{Aad:2015auj}. Similar considerations apply also for the CMS
  measurement~\cite{Khachatryan:2015oaa}.}
As a consequence, we anticipate that the more accurate NLO QCD+EW theory is
likely to be easily accommodated by the large data uncertainty, should the data
be fitted with the inclusion of EW corrections; PDFs will possibly not become
more precise.
